### 2025-03-10 paper/sections/experiments.qmd
try that again

>>>>

### Evaluation

Our key outcome of interest is how well models perform with respect to explainability (@def-explainability): to this end, we focus primarily on the plausibility and cost of faithfully generated counterfactuals at test time. To measure the cost of counterfactuals, we follow the standard convention of using distances ($\ell_1$-norm) between factuals and counterfactuals as a proxy. For plausibility, we assess how similar counterfactuals are to observed samples in the target domain. We rely on previously used distance-based metrics [@giudot]

We also assess their predictive performance using conventional performance metrics including accuracy and the F1-score.

### 2025-03-10 paper/sections/other/grids.qmd
work on paper
try that again

We begin by summarizing the high-level findings in @sec-app-grid-hl. For each of the categories, @sec-app-grid-gen to @sec-app-grid-train then present all details including the parameter grids, average performance outcomes and key evaluation metrics for the generated counterfactuals.

@@

Overall, we find that Counterfactual Training (CT) achieves it key objectives consistently across all hyperparameter settings and also broadly across datasets. We do observe strong sensitivity to certain hyperparameters, but clear and manageable patterns emerge in those cases. With respect to the underlying training data distribution, we find that outcomes for CT seem to depend to some degree on class separability.

@@

We find that CT is associated with little to no decrease in average predictive performance for our synthetic datasets: test accuracy and F1-scores decrease by at most ~1 percentage point, but generally much less (@tbl-acc-gen, @tbl-acc-pen, @tbl-acc-train).

>>>>

We begin by summarizing the high-level findings in @sec-app-grid-hl. For each of the categories, @sec-app-grid-gen to @sec-app-grid-train then present all details including the exact parameter grids, average predictive performance outcomes and key evaluation metrics for the generated counterfactuals.

### Evaluation Details

To measure predictive performance, we compute the accuracy and F1-score for all models on test data (@tbl-acc-gen, @tbl-acc-pen, @tbl-acc-train).

We evaluated

@@

Overall, we find that Counterfactual Training (CT) achieves it key objectives consistently across all hyperparameter settings and also broadly across datasets: plausibility is improved  We do observe strong sensitivity to certain hyperparameters, but clear and manageable patterns emerge in those cases. With respect to the underlying training data distribution, we find that outcomes for CT seem to depend to some degree on class separability.

@@

We find that CT is associated with little to no decrease in average predictive performance for our synthetic datasets: test accuracy and F1-scores decrease by at most ~1 percentage point, but generally much less (@tbl-acc-gen, @tbl-acc-pen, @tbl-acc-train). Variation across hyperparameters is negligible as indicated by small standard deviations for these metrics across the board.
